### 2025-09-29 YEAR2/AER210_LectureNotes/main.tex
sept 29

\end{example}
\chapter{Fluid Mechanics}
\end{document}

>>>>

    We find the Jacobian of this transformation by its inverse:
    $$
    J^{-1} = \frac{\partial(u,v,w)}{\partial(x,y,z)} = \begin{vmatrix}
        \dfrac{\partial u}{\partial x} & \dfrac{\partial u}{\partial y} & \dfrac{\partial u}{\partial z} \\
        \dfrac{\partial v}{\partial x} & \dfrac{\partial v}{\partial y} & \dfrac{\partial v}{\partial z} \\
        \dfrac{\partial w}{\partial x} & \dfrac{\partial w}{\partial y} & \dfrac{\partial w}{\partial z}
    \end{vmatrix} = \begin{vmatrix}
        y & x & 0 \\
        z & 0 & x \\
        0 & z & y
    \end{vmatrix} = xyz
    $$
    So, we can express the volume of the region in terms of $u$, $v$ and $w$:
    \begin{align*}
        V &= \iiint_R 1 \, dV = \iiint_S |J| \, du \, dv \, dw = \iiint_S \frac{1}{xyz} \, du \, dv \, dw \\
        &= \int_{25}^{49} \int_1^3 \int_1^3 \frac{1}{\sqrt{\frac{uvw}{y^2}}} \, du \, dv \, dw = \int_{25}^{49} \int_1^3 \int_1^3 \frac{y}{\sqrt{uvw}} \, du \, dv \, dw \\
        &= \int_{25}^{49} \int_1^3 \int_1^3 \frac{5}{\sqrt{uvw}} \, du \, dv \, dw = 5 \int_{25}^{49} \int_1^3 \left[ 2\sqrt{u} \right]_{u=1}^{u=3} \frac{1}{\sqrt{vw}} \, dv \, dw \\
        &= 10(\sqrt{3} - 1) \int_{25}^{49} \left[ 2\sqrt{v} \right]_{v=1}^{v=3} \frac{1}{\sqrt{w}} \, dw = 20(\sqrt{3} - 1)(\sqrt{3} - 1) \int_{25}^{49} \frac{1}{\sqrt{w}} \, dw \\
        &= 20(\sqrt{3} - 1)^2 \left[ 2\sqrt{w} \right]_{w=25}^{w=49} = 40(\sqrt{3} - 1)^2 (7 - 5) = 80(\sqrt{3} - 1)^2
    \end{align*}

\section{Vector Calculus}
\section{Line Integrals}
\begin{definition}[Line Integral of a Scalar Field]
    Consider a scalar field $f(x,y) $ and a smooth curve $C$. Then we define the line integral of $f$ along $C$ as the following Riemann sum:
    $$
        \int_C f(x,y) \, ds = \lim_{||P|| \to 0} \sum_{i=1}^n f(x_i^*, y_i^*) \Delta s_i
    $$
    where $P$ is a partition of $C$ into $n$ subintervals, $||P||$ is the norm of the partition, $(x_i^*, y_i^*)$ is a sample point in the $i$-th subinterval, and $\Delta s_i$ is the length of the $i$-th subinterval.
\end{definition}

\begin{definition}[Parametrization of a Curve]
    A smooth curve $C$ in the plane can be parametrized by:
    \begin{equation}
        \begin{cases}
            x = g(t) \\
            y = h(t)
        \end{cases} \implies \vec{r(t)} = \langle g(t), h(t) \rangle, \quad a \leq t \leq b
    \end{equation}
    where $g$ and $h$ are continuous functions with continuous derivatives on $[a,b]$.

    Under the following asummumption, we can solve the line integral using the parametrization:
    \begin{itemize}
        \item The function $f(x,y)$ is continuous on $C$.
        \item The curve $C$ is smooth and can be parametrized by $\vec{r(t)}$.
        \begin{itemize}
            \item $r'(t)$ is continuous on $[a,b]$.
            \item $r'(t) \neq 0$ for all $t \in [a,b]$. (Except possibly at the endpoints.) 
        \end{itemize}
    \end{itemize}
\end{definition}

\paragraph{Change of variable from $ds$} We can express the line integral in terms of the parameter $t$ as follows:
\begin{equation}
    \int_C f(x,y) \, ds = \int_a^b f(g(t), h(t)) \, ||\vec{r'(t)}|| \, dt
\end{equation}
where $||\vec{r'(t)}|| = \sqrt{(g'(t))^2 + (h'(t))^2}$ is the speed of the curve.

\begin{definition}[Center of Mass via Line Integral]
    Consider a wire in the plane with a density function $\rho(x,y)$ and a smooth curve $C$ parametrized by $\vec{r(t)} = \langle g(t), h(t) \rangle$, $a \leq t \leq b$. Then, the mass of the wire is given by:
    \begin{equation}
        m = \int_C \rho(x,y) \, ds = \int_a^b \rho(g(t), h(t)) \, ||\vec{r'(t)}|| \, dt
    \end{equation}
    And the center of mass $(\bar{x}, \bar{y})$ of the wire is given by:
    \begin{align}
        \bar{x} &= \frac{1}{m} \int_C x \rho(x,y) \, ds = \frac{1}{m} \int_a^b g(t) \rho(g(t), h(t)) \, ||\vec{r'(t)}|| \, dt \\
        \bar{y} &= \frac{1}{m} \int_C y \rho(x,y) \, ds = \frac{1}{m} \int_a^b h(t) \rho(g(t), h(t)) \, ||\vec{r'(t)}|| \, dt
    \end{align}
\end{definition}

\begin{example}[Center of Mass of a Semi-Circular Wire]
    Let $a$ be the radius of a semi-circular wire. We want to find the center of mass of the wire. We can parametrize the semi-circular wire as follows:
    $$
    \begin{cases}
        x = a \cos t \\
        y = a \sin t
    \end{cases} \quad 0 \leq t \leq \pi
    $$
    The density function is constant, $\rho(x,y) = \rho_0$. The mass of the wire is just the length of the wire times the density:
    $$
        m = \pi a \rho_0
    $$
    The $x$-coordinate of the center of mass is just 0 by symmetry. The $y$-coordinate of the center of mass is given by:
    \begin{align*}
        \bar{y} &= \frac{1}{m} \int_C y \rho(x,y) \, ds = \frac{1}{\pi a \rho_0} \int_0^\pi a \sin t \cdot \rho_0 \cdot a \, dt \\
        &= \frac{a^2}{\pi a} \int_0^\pi \sin t \, dt = \frac{a}{\pi} \left[ -\cos t \right]_0^\pi = \frac{2a}{\pi}
    \end{align*}
    So, the center of mass of the semi-circular wire is located at:
    $$
        \left( \bar{x}, \bar{y} \right) = \left( 0, \frac{2a}{\pi} \right)
    $$
\end{example}

\begin{definition}[Line Integral with Space Curve]
    Consider a scalar field $f(x,y,z)$ and a smooth space curve $C$. Then we define the line integral of $f$ along $C$ as the following Riemann sum:
    $$
        \int_C f(x,y,z) \, ds = \lim_{||P|| \to 0} \sum_{i=1}^n f(x_i^*, y_i^*, z_i^*) \Delta s_i
    $$
    where $P$ is a partition of $C$ into $n$ subintervals, $||P||$ is the norm of the partition, $(x_i^*, y_i^*, z_i^*)$ is a sample point in the $i$-th subinterval, and $\Delta s_i$ is the length of the $i$-th subinterval.

    A smooth space curve $C$ can be parametrized by:
    \begin{equation}
        \begin{cases}
            x = g(t) \\
            y = h(t) \\
            z = k(t)
        \end{cases} \implies \vec{r(t)} = \langle g(t), h(t), k(t) \rangle, \quad a \leq t \leq b
    \end{equation}
    where $g$, $h$ and $k$ are continuous functions with continuous derivatives on $[a,b]$.

    Under the following asummumption, we can solve the line integral using the parametrization:
    \begin{itemize}
        \item The function $f(x,y,z)$ is continuous on $C$.
        \item The curve $C$ is smooth and can be parametrized by $\vec{r(t)}$.
        \begin{itemize}
            \item $r'(t)$ is continuous on $[a,b]$.
            \item $r'(t) \neq 0$ for all $t \in [a,b]$. (Except possibly at the endpoints.) 
        \end{itemize}
    \end{itemize}

    We can express the line integral in terms of the parameter $t$ as follows:
    \begin{equation}
        \int_C f(x,y,z) \, ds = \int_a^b f(g(t), h(t), k(t)) \, ||\vec{r'(t)}|| \, dt
    \end{equation}
    where $||\vec{r'(t)}|| = \sqrt{(g'(t))^2 + (h'(t))^2 + (k'(t))^2}$ is the speed of the curve.
\end{definition}

\begin{example}[Mass of a Spring]
    Find the mass of a spring parametrized by:
    $$
    \begin{cases}
        x = 2 \cos t \\
        y = t \\
        z = 2 \sin t
    \end{cases} \quad 0 \leq t \leq 6\pi
    $$
    with a density function $\rho(x,y,z) = 2y$.

    The mass of the spring is given by:
    \begin{align*}
        m &= \int_C \rho(x,y,z) \, ds = \int_0^{6\pi} 2t \cdot \sqrt{(-2\sin t)^2 + 1^2 + (2\cos t)^2} \, dt \\
        &= \int_0^{6\pi} 2t \cdot \sqrt{4 + 1} \, dt = \sqrt{5} \int_0^{6\pi} 2t \, dt = \sqrt{5} \left[ t^2 \right]_0^{6\pi} = 36\pi^2 \sqrt{5}
    \end{align*}    
\end{example}

\begin{definition}[Line Intergral with Piecewise Smooth Curve]
    Consider a scalar field $f(x,y)$ and a piecewise smooth curve $C$ composed of $n$ smooth curves $C_1, C_2, \ldots, C_n$. Then we define the line integral of $f$ along $C$ as the sum of the line integrals along each smooth curve:
    \begin{equation}
        \int_C f(x,y) \, ds = \sum_{i=1}^n \int_{C_i} f(x,y) \, ds
    \end{equation}
\end{definition}

\begin{theorem}[Directionality of Scalar Line Integrals] The line integral of a scalar field is independent of the direction of the curve. That is, if we reverse the direction of the curve, the value of the line integral remains the same:
\begin{equation}
    \int_C f(x,y) \, ds = \int_{-C} f(x,y) \, ds
\end{equation}
\end{theorem}
\begin{proof}
    Let $C$ be a smooth curve parametrized by $\vec{r(t)} = \langle g(t), h(t) \rangle$, $a \leq t \leq b$. Then, the line integral of $f$ along $C$ is given by:
    \begin{equation}
        \int_C f(x,y) \, ds = \int_a^b f(g(t), h(t)) \, ||\vec{r'(t)}|| \, dt
    \end{equation}
    Now, consider the curve $-C$ obtained by reversing the direction of $C$. We can parametrize $-C$ by $\vec{r(-t)} = \langle g(-t), h(-t) \rangle$, $-b \leq t \leq -a$. The line integral of $f$ along $-C$ is given by:
    \begin{align*}
        \int_{-C} f(x,y) \, ds &= \int_{-b}^{-a} f(g(-t), h(-t)) \, ||\vec{r'(-t)}|| \, dt \\
        &= \int_a^b f(g(u), h(u)) \, ||\vec{r'(u)}|| \, du \quad (u = -t, du = -dt) \\
        &= \int_a^b f(g(t), h(t)) \, ||\vec{r'(t)}|| \, dt
    \end{align*}
    Thus, we have shown that:
    $$
        \int_C f(x,y) \, ds = \int_{-C} f(x,y) \, ds
    $$
\end{proof}
\chapter{Fluid Mechanics}
\end{document}

### 2025-09-29 YEAR2/CHE260_LectureNotes/main.tex
sept 29

\begin{definition}[Maxwell-Boltzmann Distribution]
    Consider a system with the $n$ particle and energy levels $\epsilon_i$. The system must first satisfy:
    $$
        \sum_i n_i = n \quad \text{and} \quad \sum_i n_i \epsilon_i = U
    $$ 
    That gives us the number of particles in a particular energy level given a temperature $T$. And this is visualized by the Maxwell-Boltzmann distribution:
    \begin{figure}[h!]
        \centering
        \includegraphics[width=0.8\textwidth]{MBGraph2.png}
        \caption{Maxwell-Boltzmann distribution of particles in energy levels}
    \end{figure}
\end{definition}
\begin{definition}[Change of Entropy for an Ideal Gas]
    The change in entropy of an ideal gas is:
    \begin{equation}
        \Delta S = s_2 - s_1 = k_B \ln\left(\frac{\Omega_2}{\Omega_1}\right) \\
    \end{equation}
    After some derivation, we have:
    \begin{equation}
        \Delta S = mR \left[ \ln\left(\frac{3}{2} \frac{T_2}{T_1}\right) + \ln\left(\frac{V_2}{V_1}\right) \right]
    \end{equation}
\end{definition}
\chapter{Heat Transfer}

>>>>

\begin{definition}[Maxwell-Boltzmann Distribution]
    Consider a system with the $n$ particle and energy levels $\epsilon_i$. The system must first satisfy:
    $$
        \sum_i n_i = n \quad \text{and} \quad \sum_i n_i \epsilon_i = U
    $$ 
    That gives us the number of particles in a particular energy level given a temperature $T$. And this is visualized by the Maxwell-Boltzmann distribution:
    \begin{figure}[h!]
        \centering
        \includegraphics[width=0.8\textwidth]{MBGraph2.png}
        \caption{Maxwell-Boltzmann distribution of particles in energy levels}
    \end{figure}
\end{definition}
\begin{definition}[Change of Entropy for an Ideal Gas]
    The change in entropy of an ideal gas is:
    \begin{equation}
        \Delta S = s_2 - s_1 = k_B \ln\left(\frac{\Omega_2}{\Omega_1}\right) \\
    \end{equation}
    After some derivation, we have:
    \begin{equation}
        \Delta S = mR \left[ \ln\left(\frac{3}{2} \frac{T_2}{T_1}\right) + \ln\left(\frac{V_2}{V_1}\right) \right]
    \end{equation}
\end{definition}

\section{The Postulate of Classical Thermodynamics}
We can now summarize thermal dynamics as a logical system.

\paragraph{Fundamental Concepts} Below are the fundamental properties of thermodynamics:
\begin{itemize}
    \item \textbf{Volume} $V$: the space occupied by the system.
    \item \textbf{Energy} $U$: the internal energy of the system.
    \item \textbf{Entropy} $S$: the measure of disorder or randomness in the system.
    \item \textbf{Mass} $m$: the amount of matter in the system.
\end{itemize}

\paragraph{Postulates} The postulates of classical thermodynamics are:
\begin{itemize}
    \item \textbf{State Postulate}: The state of a simple compressible system is completely specified by:
    \begin{equation}
        S = S(U, V, m)
    \end{equation} 
    \item \textbf{First Law of Thermodynamics}: The change in internal energy of a system is equal to the heat added to the system minus the work done by the system:
    \begin{equation}
        \Delta U = Q - W
    \end{equation}
    \item \textbf{Second Law of Thermodynamics}: The entropy of an isolated system never decreases; it either increases or remains constant for a reversible process:
    \begin{equation}
        \Delta S \geq 0
    \end{equation}
    \item \textbf{Third Law of Thermodynamics}: The entropy of a perfect crystal at absolute zero temperature is zero:
    \begin{equation}
        S = 0
    \end{equation}
    when:
    \begin{equation}
        T = 0\, \text{K} \quad \text{or} \quad \frac{\partial U}{\partial S} = 0
    \end{equation}
\end{itemize}

\chapter{Heat Transfer}

### 2025-09-29 YEAR2/ECE253_LectureNotes/main.tex
sept 29

\paragraph{Resets} We have two types of rests to reset the FF to a known state:
\begin{itemize}
    \item \textbf{Synchronous reset:} Active only on the clock edge. If reset = 1 at the triggering edge the flip‑flop is forced to the predefined state (usually 0); otherwise it captures D.
    \item \textbf{Asynchronous reset:} Active immediately when asserted, forcing the flip‑flop to the predefined state (usually 0) regardless of the clock; normal edge‑triggered operation resumes when deasserted.
\end{itemize}
\begin{definition}[Synchronous Reset D-FF]
    The Verilog implementation is as follows:
    \begin{verbatim}
    module d_ff_sync_reset(input logic D, CLK, RST,
                           output logic Q);
        always_ff @(posedge CLK) begin
            if (RST==1'b0)
                Q <= 1'b0; // Reset to 0
            else
                Q <= D; // Normal operation
        end
    endmodule
    \end{verbatim}

\end{definition}
\section{Finite State Machines (FSM)}
% FSM modeling, state diagrams, and Verilog implementation (to be expanded)

>>>>

\paragraph{Resets} We have two types of rests to reset the FF to a known state:
\begin{itemize}
    \item \textbf{Synchronous reset:} Active only on the clock edge. If reset = 1 (or resetn = 0) at the triggering edge the flip‑flop is forced to the predefined state (usually 0); otherwise it captures D.
    \item \textbf{Asynchronous reset:} Active immediately when asserted, forcing the flip‑flop to the predefined state (usually 0) regardless of the clock; normal edge‑triggered operation resumes when deasserted.
\end{itemize}
\begin{definition}[Synchronous Reset D-FF]
    The Verilog implementation is as follows:
    \begin{verbatim}
    module d_ff_sync_reset(input logic D, CLK, RSTn,
                           output logic Q);
        always_ff @(posedge CLK) begin
            if (RSTn == 1'b0)
                Q <= 1'b0; // Reset to 0
            else
                Q <= D; // Normal operation
        end
    endmodule
    \end{verbatim}

\begin{definition}[Asynchronous Reset D-FF]
    The Verilog implementation is as follows:
    \begin{verbatim}
    module d_ff_async_reset(input logic D, CLK, RSTn,
                            output logic Q);
        always_ff @(posedge CLK or negedge RSTn) begin
            if (RSTn == 1'b0)
                Q <= 1'b0; // Reset to 0
            else
                Q <= D; // Normal operation
        end
    endmodule
    \end{verbatim}
\end{definition}

\paragraph{Enable Inputs} We can add an enable input to a D-FF by adding a mux before the D input. It determines whether data is loaded on the clock edge or the old data is retained. If the enable input is high (1), the data on the D input is loaded into the flip-flop on the rising edge of the clock. If the enable input is low (0), there would be a feedback signal going into the mux, so the old data is retained.
\begin{definition}[Counters]
    To count binary numbers, we have the follow truth table:
    \begin{table}[h!]
        \centering
        \begin{tabular}{|c|c|}
            \hline
            CLK Cycle & Binary Count (Q2 Q1 Q0) \\
            \hline
            0 & 000 \\
            1 & 001 \\
            2 & 010 \\
            3 & 011 \\
            4 & 100 \\
            5 & 101 \\
            6 & 110 \\
            7 & 111 \\
            \hline
        \end{tabular}
        \caption{Truth Table for a 3-bit Counter}
        \label{tab:3_bit_counter_truth_table}
    \end{table}
    From the truth table, we can derive the following rules:
    \begin{itemize}
        \item $Q_0$ toggles every clock cycle.
        $$
        Q_0(t+1) = Q_0(t) \oplus 1
        $$
        \item $Q_1$ toggles when $Q_0$ is 1.
        $$
        Q_1(t+1) = Q_1(t) \oplus Q_0(t)
        $$
        \item $Q_2$ toggles when $Q_1$ and $Q_0$ are both 1.
        $$
        Q_2(t+1) = Q_2(t) \oplus (Q_1(t) Q_0(t))
        $$
    \end{itemize}
    We call this a TQ flip-flop, where T stands for toggle. It is done by connecting the Q output back, to a new XOR gate. The input of that XOR gate is Q and T. And the XOR gate output is connected to the D input of a D-FF. The Verilog implementation is as follows:
    \begin{verbatim}
    module counter3(input logic CLK, RSTn,
                    output logic [3:0] Q);
        always_ff @(posedge CLK or negedge RSTn) begin
            if (RSTn == 1'b0)
                Q <= 4'b0000;
            else begin
                Q[0] <= Q[0] ^ 1'b1; // Toggle Q0
                Q[1] <= Q[1] ^ Q[0]; // Toggle Q1
                Q[2] <= Q[2] ^ (Q[1] & Q[0]); // Toggle Q2
                Q[3] <= Q[3] ^ (Q[2] & Q[1] & Q[0]); // Toggle Q3
            end
        end
    endmodule
    \end{verbatim}
\end{definition}
\section{Finite State Machines (FSM)}

### 2025-09-29 YEAR2/MAT292_LectureNotes/main.tex
sept 29

>>>>

\subsection{Wronskian and Fundamental Matrix}
\begin{definition}[Functionals]
    A functional is a mapping from a vector space of functions to the real numbers. For example, consider the space of continuous functions on an interval \( [a, b] \), denoted by \( C[a, b] \). A functional \( J: C[a, b] \to \mathbb{R} \) could be defined as:
    $$
    J[f] = \int_a^b f(x) \, dx
    $$
    for any function \( f \in C[a, b] \).
\end{definition}

\paragraph{Motivation} For an $n$th-order differential equation we need $n$ linearly independent solutions to form the general solution. These solutions constitute a basis of the solution space, so we must check whether a given set of solutions is linearly independent.

\begin{definition}[Wronskian]
    Consider \( n \) functions \( x_1(t), x_2(t), \ldots, x_n(t) \) for which \( x: \mathbb{R} \to \mathbb{R}^n \) that are solutions of a $n$th-order linear differential equation on an interval \( I \). The Wronskian \( W(t) \) is defined as the determinant of the matrix whose columns are formed by the functions and their derivatives up to order \( n-1 \):
    \begin{equation}
    W(t) = \det\begin{bmatrix}
    x_1(t) & x_2(t) & \cdots & x_n(t) \\
    x_1'(t) & x_2'(t) & \cdots & x_n'(t) \\
    \vdots & \vdots & \ddots & \vdots \\
    x_1^{(n-1)}(t) & x_2^{(n-1)}(t) & \cdots & x_n^{(n-1)}(t)
    \end{bmatrix} 
    \end{equation}
    Iff there exist \( t \) in interval \( I \) such that \( W(t) \neq 0 \), then the functions \( x_1(t), x_2(t), \ldots, x_n(t) \) are linearly independent on \( I \).

\begin{example}
    Let $f(x) = \cos (mx)$ and $g(x) = \sin (mx)$ for some constant $m \neq 0$. We compute the Wronskian:
    \begin{align*}
    W(f, g)(x) &= \det\begin{bmatrix}
    \cos(mx) & \sin(mx) \\
    -m\sin(mx) & m\cos(mx)
    \end{bmatrix} \\
    &= m\cos^2(mx) + m\sin^2(mx) \\
    &= m(\cos^2(mx) + \sin^2(mx)) \\
    &= m \neq 0
    \end{align*}
    Since \( W(f, g)(x) \neq 0 \) for all \( x \), the functions \( \cos(mx) \) and \( \sin(mx) \) are linearly independent on \( \mathbb{R} \).
\end{example}
\paragraph{Trading order for dimension} Consider the $n$th-order ODE:
$$a_n(t)x^{(n)} + a_{n-1}(t)x^{(n-1)} + \ldots + a_1(t)x' + a_0(t)x = g(t)$$
where \( a_n(t) \neq 0 \) on an interval \( I \). We can convert this into a system of first-order ODEs by defining:
$$\begin{cases}
x_1 = x \\
x_2 = x' \\
x_3 = x'' \\
\vdots \\
x_n = x^{(n-1)}
\end{cases}$$
Then, we have solution $x: \mathbb{R} \to \mathbb{R}^n$ given by:
$$
x = v_1 e^{\lambda_1 t} + v_2 e^{\lambda_2 t} + \ldots + v_n e^{\lambda_n t}
$$
where \( v_i \) are the eigenvectors and \( \lambda_i \) are the eigenvalues of the corresponding matrix \( A \). Then, we can define the Wronskian for this system as:
\begin{equation}
W(v_1e^{\lambda_1 t}, v_2 e^{\lambda_2 t}, \ldots, v_n e^{\lambda_n t}) = 
\begin{vmatrix}
v_1 e^{\lambda_1 t} & v_2 e^{\lambda_2 t} & \ldots & v_n e^{\lambda_n t} 
\end{vmatrix} = e^{(\lambda_1 + \lambda_2 + \ldots + \lambda_n)t} \begin{vmatrix}
v_1 & v_2 & \ldots & v_n 
\end{vmatrix}
\end{equation}
Which the first term is never zero, so the Wronskian is non-zero iff the eigenvectors are linearly independent. The basis $\{v_1 e^{\lambda_1 t}, v_2 e^{\lambda_2 t}, \ldots, v_n e^{\lambda_n t}\}$ is a fundamental called the fundamental set of solutions.

\paragraph{General Planer Systems} Consider the system of ODEs:
$$\textbf{x}' = A\textbf{x}$$
where \( A \in \mathbb{R}^{2 \times 2} \).

### 2025-09-29 YEAR2/PHY293_LectureNotes/main.tex
sept 29

\begin{example}[Vibrating String]
    Consider a string in the $xy$-plane, stretched along the x-axis with tension \( \tau \) and linear mass density \( \mu \). Now, consider at the end of the string at $x$ makes a small angle \( \theta \) with the x-axis. Then the components of the tension are:
    $$
    \begin{cases}
        \tau_x &= \tau \cos(\theta) \approx \tau \\
        \tau_y &= \tau \sin(\theta) \approx \tau \frac{\partial y}{\partial x}
    \end{cases}
    $$
    Now, consider a small segment of the string between \( x \) and \( x + \delta x \). The vetical force on the other end is given, via linear approximation, by:
    $$    
        \tau_y(x + \delta x) \approx \tau_y(x) \left[ \frac{\partial y}{\partial x}_x + \frac{\partial^2 y}{\partial x^2} \delta x \right]
    $$
    So the net force is given by:
    $$
        F_y = \tau_y(x + \delta x) - \tau_y(x) = \tau \frac{\partial^2 y}{\partial x^2} \delta x
    $$
    Becuase $F=ma$, we have:
    $$
        \tau \frac{\partial^2 y}{\partial x^2} \delta x = dm \frac{\partial^2 y}{\partial t^2}
    $$
    where \( dm = \mu \delta x \) is the mass of the small segment. Thus, we have:
    $$
        \frac{\partial^2 y}{\partial t^2} - v^2 \frac{\partial^2 y}{\partial x^2} = 0
    $$
    where \( v = \sqrt{\frac{\tau}{\mu}} \) is the wave speed on the string. This is the one-dimensional wave equation.

\subsection{The Wave Equation}

\begin{definition}[The Wave Equation]
    The wave equation is a second-order linear partial differential equation that describes the propagation of waves (sound, EM, water). In one dimension, it is:
    \begin{equation}
        \frac{\partial^2 u}{\partial t^2} - c^2 \frac{\partial^2 u}{\partial x^2} = 0
    \end{equation}
    where $u(x,t)$ is the disturbance, $c$ is the wave speed, $x$ is position, and $t$ is time.
\end{definition}

\chapter{Modern Physics}
\end{document}

>>>>

\begin{example}[Vibrating String]
    Consider a string in the $xy$-plane, stretched along the x-axis with tension \( \tau \) and linear mass density \( \mu \). Now, consider at the end of the string at $x$ makes a small angle \( \theta \) with the x-axis. Then the components of the tension are:
    $$
    \begin{cases}
        \tau_x &= \tau \cos(\theta) \approx \tau \\
        \tau_y &= \tau \sin(\theta) \approx \tau \frac{\partial y}{\partial x}
    \end{cases}
    $$
    Now, consider a small segment of the string between \( x \) and \( x + \delta x \). The vetical force on the other end is given, via linear approximation, by:
    $$    
        \tau_y(x + \delta x) \approx \tau_y(x) \left[ \frac{\partial y}{\partial x}_x + \frac{\partial^2 y}{\partial x^2} \delta x \right]
    $$
    So the net force is given by:
    $$
        F_y = \tau_y(x + \delta x) - \tau_y(x) = \tau \frac{\partial^2 y}{\partial x^2} \delta x
    $$
    Because \( F=ma \), we have:
    $$
        \tau \frac{\partial^2 y}{\partial x^2} \delta x = \text{d}m \frac{\partial^2 y}{\partial t^2}
    $$
    where \( \text{d}m = \mu \delta x \) is the mass of the small segment. Thus, we have:
    $$
        \frac{\partial^2 y}{\partial t^2} - v^2 \frac{\partial^2 y}{\partial x^2} = 0
    $$
    where \( v = \sqrt{\frac{\tau}{\mu}} \) is the wave speed on the string. This is the one-dimensional wave equation.

\end{example}
\begin{definition}[The Wave Equation]
    The wave equation is a second-order linear partial differential equation that describes the propagation of waves (sound, EM, water) through a medium. In one dimension, it is given by:
    \begin{equation}
        \frac{\partial^2 u}{\partial t^2} - c^2 \frac{\partial^2 u}{\partial x^2} = 0
    \end{equation}
    where $u(x,t)$ is the disturbance, $c$ is the wave speed, $x$ is position, and $t$ is time.

    Or in general:
    \begin{equation}
        \frac{\partial^2 u}{\partial t^2} - c^2 \nabla^2 u = 0
    \end{equation}
    where \( \nabla^2 \) is the Laplacian operator.
\end{definition}

\begin{definition}[Energy of a String]
    The kinetic energy of a string segment of mass \( \text{d}m = \mu \text{d}x \) is given by:
    \begin{equation}
        \text{d}K = \frac{1}{2} \text{d}m v_y^2 = \frac{1}{2} \mu \text{d}x \left(\frac{\partial y}{\partial t}\right)^2
    \end{equation}
    The potential energy stored in the string due to its tension is given by:
    \begin{equation}
        \text{d}U = \frac{1}{2} \tau (\delta s - \delta x) 
    \end{equation}
    where \( \delta s \) is the actual length of the string segment and \( \delta x \) is the horizontal length. For small displacements, we can approximate:
    $$
        \delta s \approx \delta x \left(1 + \frac{1}{2}\left(\frac{\partial y}{\partial x}\right)^2\right)
    $$
    Thus, the potential energy becomes:
    \begin{equation}
        \text{d}U \approx \frac{1}{2} \tau \delta x \left(\frac{\partial y}{\partial x}\right)^2
    \end{equation}
    The total energy \( E \) of the string segment is the sum of its kinetic and potential energies:
    \begin{equation}
        \text{d}E = \frac{1}{2} \mu \text{d}x \left(\frac{\partial y}{\partial t}\right)^2 + \frac{1}{2} \tau \text{d}x \left(\frac{\partial y}{\partial x}\right)^2
    \end{equation}
    For a sinusoidal wave \( y(x,t) = A \cos(kx - \omega t) \), the change in energy is:
    \begin{equation}
        \text{d}E = \frac{1}{2} \mu A^2 \omega^2 \sin^2(kx - \omega t) \text{d}x + \frac{1}{2} \tau A^2 k^2 \sin^2(kx - \omega t) \text{d}x
    \end{equation}
    Integrating over one wavelength \( \lambda \), we find the total energy in one wavelength:
    \begin{equation}
        E = \int_0^\lambda \text{d}E = \frac{1}{4} \mu A^2 \omega^2 \lambda + \frac{1}{4} \tau A^2 k^2 \lambda
    \end{equation}

\begin{definition}[Power Transmitted by a String]
   The power carried by a transverse wave could be calculated from $P=Fv$. The transverse force on the string is given by:
    \begin{equation}
        P(x,t) = A^2 \mu v \omega^2 \sin^2(kx - \omega t + \phi_0)
    \end{equation}
    The average power is then given by:
    \begin{equation}
        \bar{P} = \frac{1}{2} A^2 \mu v \omega^2 = \frac{1}{2} \mu A^2 \omega^2 \frac{\lambda}{T}
    \end{equation}
    And the maximum power is given by:
    \begin{equation}
        P_{max} = A^2 \mu v \omega^2
    \end{equation}
\end{definition}

\paragraph{Speed of Sound} The speed of sound waves through various meida is as follows:
\begin{itemize}
    \item In ideal gas, it is given by:
    \begin{equation}
        v = \sqrt{\frac{\gamma RT}{m}} = \sqrt{\frac{\gamma k_B T}{m}}
    \end{equation}
    where \( \gamma \) is the adiabatic index, \( R \) is the universal gas constant, \( k_B \) is the Boltzmann constant, \( T \) is the absolute temperature, and \( m \) is the molar mass of the gas. $\gamma$ varies based on the type of gas:
    \begin{itemize}
        \item Monatomic gases (e.g., helium, neon): \( \gamma = \frac{5}{3} \)
        \item Diatomic gases (e.g., nitrogen, oxygen): \( \gamma = \frac{7}{5} \)
        \item Polyatomic gases (e.g., carbon dioxide, methane): \( \gamma \) is typically between 1.3 and 1.4
    \end{itemize}
    \item In a liquid, it is given by:
    \begin{equation}
        v = \sqrt{\frac{B_{\text{adiabatic}}}{m}} = \sqrt{\frac{B}{\rho}}
    \end{equation}
    where \(B_{\text{adiabatic}}\) is the adiabatic bulk modulus, \( B \) is the bulk modulus, \( \rho \) is the density of the liquid and \( m \) is the molar mass of the liquid.
    \item In a solid, it is given by:
    \begin{equation}
        v = \sqrt{\frac{E}{\rho}} 
    \end{equation}
    where \( E \) is the Young's modulus and \( \rho \) is the density of the solid. It could also be given by:
    $$
        v = \sqrt{\frac{B}{\rho}}
    $$
    where \( B \) is the bulk modulus.
\end{itemize}

\begin{definition}[Impedance]
    Similar to electrical impedance, we also have acoustic impedance and mechanical, they measure how much a structure recisits motion when subjected to a harmonic force. They are defined as follows:
    \begin{itemize}
        \item Electical impedance \( Z_e \):
        \begin{equation}
            Z_e = \sqrt{R^2 + \left(\omega L - \frac{1}{\omega C}\right)^2}
        \end{equation}
        \item Acoustic impedance \( Z_a \), it is the ratio of acoustic pressure to particle velocity:
        \begin{equation}
            Z_a = \frac{p}{v} = \rho v = \sqrt{B\rho}
        \end{equation}
        \item Mechanical impedance \( Z_m \):
        \begin{equation}
            Z_m = v\mu = \sqrt{\tau\mu}
        \end{equation}
    \end{itemize}
\end{definition}

\paragraph{Wave boundry Conditions} When a wave encounters a boundary between two different media, part of the wave is reflected back into the original medium, and part is transmitted into the new medium. This is constrined by the following boundary conditions:
\begin{itemize}
    \item The displacement of the wave must be continuous across the boundary:
    \begin{equation}
        y_1(x_b, t) = y_2(x_b, t) = y_3(x_b, t)
    \end{equation}
    \item The derivative of the displacement with respect to position must also be continuous across the boundary:
    \begin{equation}
        \frac{\partial y_1}{\partial x}\bigg|_{x_b} = \frac{\partial y_2}{\partial x}\bigg|_{x_b} = \frac{\partial y_3}{\partial x}\bigg|_{x_b}
    \end{equation}
\end{itemize}
and then from these conditions, we can derive the amplitudes of the reflected and transmitted waves:
\begin{equation}
    A_i = A_r + A_t
\end{equation}
and 
\begin{equation}
    A_i Z_1 = A_r Z_1 + A_t Z_2
\end{equation}
From these two equations, we can solve for the reflection and transmission coefficients:
\begin{equation}
    R = \frac{A_r}{A_i} = \frac{Z_2 - Z_1}{Z_2 + Z_1}
\end{equation}
and
\begin{equation}
    T = \frac{A_t}{A_i} = \frac{2Z_2}{Z_2 + Z_1}
\end{equation}
\chapter{Modern Physics}
\end{document}
